Region-network hierarchical sparsity priors for\\ high-dimensional inference in brain imaging

### 2016-01-26 paper/paper.tex
first complete results

\section{Experimental Results}
\paragraph{Serial versus parallel structure discovery and classification.}
\paragraph{Serial versus parallel structure discovery and classification.}
\paragraph{Serial versus parallel structure discovery and classification.}

@@

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/reg_net_ratio.pdf}
\end{centering}
\vspace{-0.6cm}
\caption{\textbf{Support recovery as a function of
region and network emphasis.}
The relative impact of the region and network priors
on model selection
is systematically varied against each other.
This region-network ratio (\textit{upper row}) weighted voxel groups
to priviledge sparse models in function space
that acknowledge known brain region neighborhoods
(\textit{left columns}) or
known brain networks compartments
(\textit{right columns}).
Among the 18 classes, the model weights are shown for the
tasks (\textit{from top to bottom}): tongue movement, listening stories,
taking somebody else's perspective (ToM, "theory of mind"),
as well as
viewing locations and tools.
The 18-class out-of-sample accuracy is indicated
on the \textit{bottom} and
the class-wise mean neural activty
in the \textit{rightmost column}.
%
Different emphasis on regions versus networks
in hierarchical structured sparsity can
yield essentially similar model performance.
%
Priviledging region versus network structure during model selection
recovers complementary aspects of the brain activity pattern.
%
Equal region and network emphasis yields more dispersed,
less interpretable predictive model choices.
}
\label{fig_ratio}
\end{figure}

\paragraph{Sample complexity of na\"ive versus informed sparse model selection.}
Subsequently, the sample complexity of
$\ell_1$-penalized and hierarchical-tree-penalized logistic regression
were quantitatively compared (Figure \ref{fig_dataratio}).
Region-network priors should bias model selection towards more
neurobiologically plausible classification estimators.
This should yield better out-of-sample generalization and
support recovery than
$\ell_1$-constrained logistic regression na\"ive to neurobiology
in the data-scarce and data-rich scenarios.
%
The HCP task data with examples from 18 classes were first divided into
90\% of training set (i.e., 7584 neural activity maps) and
10\% of test set (i.e., 842 maps).
Both learning algorithms were fitted based on the
training set at different subsampling fractions:
20\% (1516 maps),
40\% (3033 maps),
60\% (4550 maps),
80\% (6067 maps), and
100\% (7584 maps).
The stratified and shuffled training data were submitted to
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.
%
Three observation have been made.
In the data-scarce scenario (i.e., 1/5 of actual training data),
hierarchical tree sparsity achieved the biggest advantage
in out-of-sample performance by 5,11\% as well as
better support recovery with weight maps already much closer
to the training data average.
In the case of scarce training data, which is typical for the brain imaging domain,
regularization by region-network priors indeed allowed for
more effective extraction of classification-relevant structure
from the neural activity scans.
%
Across scenarios,
the weight maps from ordinary logistic regression exhibit
higher variance and many more zero coefficients
than hierarchical tree logistic regression.
Given the usually high multicollinearity in neuroimaging data,
this observation is likely to reflect instable selection of
representatives among class-responsive predictor groups
due to the $\ell_1$-norm penalization.
%
In the data-rich scenario (i.e., entire training data used for model fitting),
neurobiologically informed logistic regression
profits more from the increased information quantities than
neurobiologically naive logistic regression.
That is, the region-network priors actually further enhance the similarity
to the weight maps even in abundant input data.
This was the case although
the maximal classification performance of $\approx$90\% has already
been reached with small training data fractions by the structured estimator.
In contrast, 
the unstructured estimator reached this generalization performance
only with bigger input data quantities.

\paragraph{Support recovery as a function of region and network emphasis.}

>>>>

\section{Experimental Results}

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/sparsities.pdf}
\end{centering}
\vspace{-0.6cm}
\caption{\textbf{Model performance across sparsity priors.}
Compares the performance of 4 logistic regression estimators
with different structured and unstructered sparsity terms
in classifying neural activity from 18 psychological tasks.
The class-wise precision and recall metrics
were obtained on the test set.
%
Unstructured $\ell_1$-penalized logistic regression (\textit{upper left})
imposed a minimum of relevant brain voxels without
assuming special structure.
Structured $\ell_1/\ell_2$ block norm with additional $\ell_1$ term
(\textit{lower left})
imposed region compartments, but na\"ive to network structure.
Structured trace-norm penalization (\textit{lower tight})
imposed low-rank structure
with sparsity of network patterns, but na\"ive to region structure.
Structured $\ell_1/\ell_2$ block norm with a hierarchy of
both region and network priors (\textit{upper tight})
exhibited the best out-of-sample performance.
%
Consequently, a priori knowledge of both region and network neighborhoods
was the most beneficial structured sparsity penalty.
}
\label{fig_sparsities}
\end{figure}

\paragraph{Benchmarking hierarchical tree sparsity against
common sparsity penalties.}
Hierarchical region-network priors have been systematically
evaluated against other popular choices of sparse classification algorithms
in an 18-class scenario
(Figure \ref{fig_sparsities}).
%
Logistic regression with $\ell_1/\ell_2$ block norm penalization
incorporated a hierarchy of previously known region and network neighborhoods
for a neurobiological bias of the statistical estimation.
%
Vanilla logistic regression with $\ell_1$-penalization
does not assume any previously known special structure.
This classification estimator embraces a vision of neural activity structure
that expectes a minimum of
topographically and functionally independent brain voxel to be relevant.
%
Logistic regression with sparse group sparsity
imposes a structured $\ell_1/\ell_2$ block norm with additional $\ell_1$ term
with a known atlas of region voxel groups onto the statistical estimation process.
This supervised estimator shrinks and selects the coefficients
of topographically compact voxel groups expected to be relevant together.
%
Logistic regression with trace-norm penalization
imposed low-rank structure.
This supervised classification algorithm
expected a minimum of unknown network patterns
to be relevant.
%
The stratified and shuffled training data were submitted to
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.
%
Hierarchial tree sparsity demonstrated the best generalization
in distinguihsing unseen neural activity maps from 18 tasks
(89.67\%, mean recall XX.YY, mean precision XX.YY).
It was closely followed by logistic regression with trace-norm
regularization that is structured but not neurobiologically informed
(89.43\%, mean recall XX.YY, mean precision XX.YY).
Lasso featured an average performance comparing to the other sparse estimators
(88.60\%, mean recall XX.YY, mean precision XX.YY)
Introduced a priori knowledge of brain region compartments
by sparse group sparsity performed worst
(87.89\%, mean recall XX.YY, mean precision XX.YY).
%
In sum,
biasing sparse model selection by domain knowledge of region-network hierarchies
outperformed other types of frequently used sparse penalization techniques.

@@

\paragraph{Sample complexity of na\"ive versus informed sparse model selection.}
Subsequently, the sample complexity of
$\ell_1$-penalized and hierarchical-tree-penalized logistic regression
were quantitatively compared (Figure \ref{fig_dataratio}).
Region-network priors should bias model selection towards more
neurobiologically plausible classification estimators.
This should yield better out-of-sample generalization and
support recovery than
$\ell_1$-constrained logistic regression na\"ive to neurobiology
in the data-scarce and data-rich scenarios.
%
The HCP task data with examples from 18 classes were first divided into
90\% of training set (i.e., 7584 neural activity maps) and
10\% of test set (i.e., 842 maps).
Both learning algorithms were fitted based on the
training set at different subsampling fractions:
20\% (1516 maps),
40\% (3033 maps),
60\% (4550 maps),
80\% (6067 maps), and
100\% (7584 maps).
%
The stratified and shuffled training data were submitted to
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.
%
Three observation have been made.
In the data-scarce scenario (i.e., 1/5 of actual training data),
hierarchical tree sparsity achieved the biggest advantage
in out-of-sample performance by 5,11\% as well as
better support recovery with weight maps already much closer
to the training data average.
In the case of scarce training data, which is typical for the brain imaging domain,
regularization by region-network priors indeed allowed for
more effective extraction of classification-relevant structure
from the neural activity scans.
%
Across scenarios,
the weight maps from ordinary logistic regression exhibit
higher variance and many more zero coefficients
than hierarchical tree logistic regression.
Given the usually high multicollinearity in neuroimaging data,
this observation is likely to reflect instable selection of
representatives among class-responsive predictor groups
due to the $\ell_1$-norm penalization.
%
In the data-rich scenario (i.e., entire training data used for model fitting),
neurobiologically informed logistic regression
profits more from the increased information quantities than
neurobiologically naive logistic regression.
That is, the region-network priors actually further enhance the similarity
to the weight maps even in abundant input data.
This was the case although
the maximal classification performance of $\approx$90\% has already
been reached with small training data fractions by the structured estimator.
In contrast, 
the unstructured estimator reached this generalization performance
only with bigger input data quantities.

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/reg_net_ratio.pdf}
\end{centering}
\vspace{-0.6cm}
\caption{\textbf{Support recovery as a function of
region and network emphasis.}
The relative impact of the region and network priors
on model selection
is systematically varied against each other.
This region-network ratio (\textit{upper row}) weighted voxel groups
to priviledge sparse models in function space
that acknowledge known brain region neighborhoods
(\textit{left columns}) or
known brain networks compartments
(\textit{right columns}).
Among the 18 classes, the model weights are shown for the
tasks (\textit{from top to bottom}): tongue movement, listening stories,
taking somebody else's perspective (ToM, "theory of mind"),
as well as
viewing locations and tools.
The 18-class out-of-sample accuracy is indicated
on the \textit{bottom} and
the class-wise mean neural activty
in the \textit{rightmost column}.
%
Different emphasis on regions versus networks
in hierarchical structured sparsity can
yield essentially similar model performance.
%
Priviledging region versus network structure during model selection
recovers complementary aspects of the brain activity pattern.
%
Equal region and network emphasis yields more dispersed,
less interpretable predictive model choices.
}
\label{fig_regnetratio}
\end{figure}

\paragraph{Support recovery as a function of region and network emphasis.}
Finally, the relative importance of the
region and network priors within the hierarchical tree prior
was quantified (Figure \ref{fig_regnetratio}).
The $\eta_g$ group of region priors was multiplied with a
region-network ratio, while the
$\eta_g$ group of network priors was biased by the corresponding
network-region ratio. A region-network ratio of 3, for instance,
increased the relative importance of known region structure
by multiplying \frac{3}{1} to the
$\eta_g$ factor of all region groups
and multiplying
\frac{1}{3} to all network groups.
The data splitting cross-validation scheme was identical to the
above modelling experiments.
%
As the most important observation,
a range between region-dominant and network-dominant structured penalties
yields quantitatively almost identical generalization to new data
but qualitatively different decision functions manifested in the weight maps
(Figure \ref{fig_regnetratio}, 2. and 4. column).
Classification models with many zero coefficients but high absolute
coefficients in either region compartments or network compartments
can similarly extrapolation to unseen neural activity maps.
Second,
these perform similar to equilibrated region-network priors
that set less voxel coefficients to zero and spread the
probability mass with lower absolute coefficients across the whole brain
(Figure \ref{fig_regnetratix}, 3. column in the middle).
Third,
overly strong emphasis on either level of the hierarchical prior
can yield the neurobiologically informative maps
of the most necessary region or network structure for
statistically significant out-of-sample performance
(Figure \ref{fig_regnetratix}, leftmost and rightmost columns).
%
In sum,
stratifying the hierarchical tree penalty between region and network emphasis
suggests that \textit{class-specific region-network weights}
might offer more performant and more interpretable classification models
in the future.
